Six ways to understand monads\\ \small{Experiences from the functional programming course at TU Delft}

### 2012-11-01 latex/paper.tex
Remove references to restrict to 4 pages

Functional programming is arguably one of the oldest programming paradigms. In it is
purest form, it is based on a minimalistic theoretical background
($\lambda$-calculus~\cite{Baren84}). Relatively recent
formulations~\cite{Meije91, Wadle93} also introduced concepts from category
theory. While the every day use of functional programming does not necessarily
require the programmer to be aware of the theory, understanding it usually leads
to more elegant algorithmic solutions. Functional programming teaching is also
associated with advanced type systems; indeed, the flagship functional
programming languages (Haskell and the {\sc ml} family) both offer very
sophisticated support for types. Consequently, teaching the full spectrum of
functional programming theory and techniques would have been impossible in a
half-semester course; instead we decided to focus on practical aspects of data
processing, transformations and state representation using functional
techniques.

@@

Coding sprints or ``hackathons'' have long been used by open source software
projects to speed up development and increase participation of community
members~\cite{Adams11}. In a typical sprint, project members set measurable
targets and work intensively in a tight deadline to deliver them. To ensure the
timely delivery of the student projects, we organized an one day coding sprint,
during which the students were requested to i) set explicit targets ii) write
code for 7 hours with a short break iii) present a 5 minute demo of their work
at the end of the day. All project teams where expected to work in the same
room, while communication among the teams was facilitated by the lab
demonstrator, to avoid unnecessary distractions.

>>>>

Functional programming is arguably one of the oldest programming paradigms. In it is
purest form, it is based on a minimalistic theoretical background
($\lambda$-calculus). Relatively recent
formulations~\cite{Meije91, Wadle93} also introduced concepts from category
theory. While the every day use of functional programming does not necessarily
require the programmer to be aware of the theory, understanding it usually leads
to more elegant algorithmic solutions. Functional programming teaching is also
associated with advanced type systems; indeed, the flagship functional
programming languages (Haskell and the {\sc ml} family) both offer very
sophisticated support for types. Consequently, teaching the full spectrum of
functional programming theory and techniques would have been impossible in a
half-semester course; instead we decided to focus on practical aspects of data
processing, transformations and state representation using functional
techniques.

@@

Coding sprints or ``hackathons'' have long been used by open source software
projects to speed up development and increase participation of community
members. In a typical sprint, project members set measurable
targets and work intensively in a tight deadline to deliver them. To ensure the
timely delivery of the student projects, we organized an one day coding sprint,
during which the students were requested to i) set explicit targets ii) write
code for 7 hours with a short break iii) present a 5 minute demo of their work
at the end of the day. All project teams where expected to work in the same
room, while communication among the teams was facilitated by the lab
demonstrator, to avoid unnecessary distractions.
